Spring haul-out behavior of seals in the Bering and Chukchi seas

### 2024-02-18 analysis/paper/Letter_Response_to_Reviewers.qmd

> ***Line 227:** Seals can just as happily be on foraging trips within the pack
> and marginal sea ice as in the open water, depending on the species.
> Suggesting that foraging trips are associated with ice-free conditions is
> misleading.*

@@

::: author-response
**\[author response\]** We have removed 'e.g., foraging trips' from this
sentence so as not to mislead the reader.
:::

> ***Line 229-230:** This is your primary reason for not including sea ice
> geography in your analysis: you simply lacked geolocation of much of your
> haulout behavior data.*
>
> *Please help the reader understand how coarse your tracking data was for the
> flipper tagged animals and what proportion of each species was tracked with
> these flipper tags. Please also help the reader understand how the tracking
> data relates geographically to regions that may be subject to aerial survey
> efforts (i.e. are your tracking data collected from across the regions that
> may be surveyed during range-wide spring aerial surveys?).*

::: author-response
**\[author response\]** We have re-organized this paragraph to, hopefully,
better communicate our choice not to explore sea-ice related covariates. As the
reviewer suggests, we have limited geo-location data for in-water periods of our
haul-out behavior data -- we know the seal was in the water but don't have any
estimates of location when only a flipper tag is attached. By leading with this
statement, we hope that communicates our primary rationale. But, the other
rationales are also important and we still include them. We have also moved this
section to the 'Discussion' in response to other reviewer comments.

@@

We have also added a figure 1 depicting a spatial representation of recent
aerial survey effort (which reflects estimated species range for bearded,
ribbon, and spotted seals in US waters). This should better communicate to the
reader how the spatial distribution of our haul-out behavior overlaps with the
species range and survey effort.
:::

> ***Line 229-230:** Please help the reader understand why you have not included
> a basic metric of sea ice geography, such as distance from marginal ice edge
> (from the NIC's daily Marginal Ice Zone charts), which is a metric that would
> be available for all data collected during an aerial survey.*

@@

::: author-response
**\[author response\]** We have revised the text in this section to try to
provide a clearer understanding of our reasoning. Also, note this has been moved
to the Discussion section.

The text now reads:

Notably missing from our list of haul-out model explanatory variables is any
spatial-temporal representation of sea ice concentration, area, or extent. This
may seem counterintuitive when modeling the haul-out behavior of seal species
with such a close association to sea ice; seals haul out in the presence of sea
ice, and we could assess the local concentration of sea ice during these events
(see [@olnes2020]). This, however, expands the scope of our analysis into the
realm of habitat selection and many of our deployments consisted of a single
device attached to the rear flipper of the seal which meant we only received
locations when seals were hauled out on sea-ice, limiting our ability to fully
evaluate fine-scale habitat preferences related to sea-ice. Insight into how
seals use and interact with sea-ice during an extended period when the
availability and characteristics of sea-ice is rapidly changing is important
[@breed2018; @cameron2018] but ancillary to the focus of this analysis -- and,
in the end, not possible given key limitations of our data. Additionally, our
study was limited to the spring season when seal haul-out behaviors are strongly
influenced by pupping, nursing, breeding behavior, and molt. These drivers are
likely more influential on haul-out behavior than sea-ice concentration.
Crawford et al. [-@crawford2019] compared haul-out probability models for ringed
seals and found those that only included season (and not sea-ice concentration)
were the most parsimonious. For all of these reasons, we have elected not to use
sea ice concentration _as a predictor for haul-out probability_ in the present
study.
:::

> ***Table 2:** It is a bit misleading to indicate that the NARR variables were
> sourced from NARR alone. Their values were interpolated onto a geogrpahic
> location that was widely variable in precision and that precision relates to
> the both the tag attachment (glued on or flipper attached) and the seal
> behavior (in water, hauled out).*
>
> *Indicating that your "Northing" variable was from the "Bio-logger" is also
> misleading because your locations was determined by either the Doppler
> inflection point time, satellite orbitography and a model of the Earth (the
> Argos relocation method) or by reprocessing of GNSS signals that were logged
> and then transmitted via the Argos system as a data message. Really, neither
> of these were determined by the "Bio-logger", rather they were determined by a
> very complex system of geolocation.*

> *Your readers may be better served by not including this "Source" field, given
> that you describe how you obtained or estimated these variable values in your
> methods text. Offering an over-simplified "Source" in the table obscures your
> actual methods.*

@@

::: author-response
**\[author response\]** We appreciate the reviewer's thoughts on this and agree
with that removing the 'Source' column better serves the readers.
:::

> ***Line 244:** Neither your telemetry data nor aerial surveys can distinguish
> "basking" as a behavioral state. You are measuring "hauled out" behavior and
> aerial surveys observe "hauled out" seals. I encourage you to use these terms
> and not mix in ideas of "basking", "nursing", "breeding", "taking refuge from
> marine predators" etc.*

@@

::: author-response
**\[author response\]** As previously mentioned, we've addressed this and
consistently use "hauled out" and not "basking"
:::

> ***Line 254:** It is nice that you have sufficient data to do this. However,
> the reason for having separate models for each species is because you expect
> different behaviors from each specifies and because you expect that aerial
> efforts will be able to distinguish sea species and be able to apply an
> availability correction factor for each species. Please correct me if I am
> wrong, else, please keep your reader focused on the value that you are
> bringing to the analysis here in the Methods section and the good fortune that
> you have in the sample size, which should be noted in the Results and
> Discussion section.*

@@

::: author-response
**\[author response\]** We have rephrased this sentence to emphasize that we fit
each species separately because a) we expected differing behaviors across
species and b) fitting a single model was not possible with such a large data
set (even with the advantages provided by the glmmLDTS package).
:::

> ***Line 263:** Please clarify what is intended by the notation "only -99% or
> our observations".*

@@

::: author-response
**\[author response\]** This is an error in the code that has been fixed.
:::

> ***Line 263:** Also, please consider using consistent terms for "behavior
> records".  Perhaps it could be fair to term your telemetered haulout status as
> an "observation", however, it seems more intuitive to reference this a
> "record"*

@@

::: author-response
**\[author response\]** We have made edits throughout the manuscript to use
consistent terms for 'behavior records'
:::

> ***Line 263:** Please help your reader understand why you have not adopted use
> of a classified variable for each telemetered behavior interval whereby it
> could be either "wet" or "dry" based on a threshold proportion of readings
> reported as "dry".  This has two advantages: (1) were this enabled through
> processing on-board the tags, it would greatly increase your data compression
> and collection; (2) it may enable greater analytical power in your modeling
> effort.*

@@

::: author-response
**\[author response\]** We provide fairly clear and detailed information on the
haul-out behavior records, how they are recorded and transmitted in the methods
section. We have further edited this to make sure it is clear that these are
standard settings within the Wildlife Computers tag platform -- we do not have
direct control over the details for on-board processing or transmission. As
stated in the text, for this analysis, the linear predictor was modeled on the
logit scale as a Bernoulli response where tags are either 'wet' or 'dry' for a
given hour based on whether the tag was greater than 50% dry. This is a
classified variable as suggested. The difference is that we conducted this
classification post-transmission instead of on-board the tag. We feel the
transmission of various percent-dry values for a given hour provide more
flexibility in the types of analyses and insights that can be gained from
deployments.
:::

> ***Line 265-272:** Please seek input from a competent biometrician. I am not
> able to offer any guidance on this other than to muse why you are pushing to
> estimate six parameters for a time of day covariate when your seals are not
> known to keep track of time.*
>
> *It would be reasonable to assume that seals respond to insolation or the
> angular height of the sun in its daily arc.  Please help your reader
> understand why you have sought to model a complex representation of hourly
> time when you could have pulled insolation from NARR or calculated sun angular
> altitude (perhaps as a linear proportion of the day's maximum sun angular
> altitude) from standard astronomical functions.*
>
> *Perhaps your motivation is to provide plots that appear intuitive with hours
> on the y-axis, day of year on the x-axis and haulout proportion as a "heat
> map".  However beautiful these plot are, I wonder if your goal for improving
> precision and reducing bias when estimating availability during an aerial
> survey would not be better served by  the alternative and simpler variables.*

::: author-response
**\[author response\]** We appreciate the suggestion and find the possibility of
using insolation values from reanalysis products as a covariate intriguing. As
the reviewer suggests, such an approach could provide a less complicated and
more parsimonious model than the approach we have taken with the Fourier series
as a cyclic representation of hour-of day. We have taken the time to do some
additional analysis and methodological exploration and provide additional
details in Supplement 2. Note, our initial exploration of this approach resulted
in lower AIC values for the Fourier series approach presented in the analysis.
While promising and worth even further exploration, we feel the approach taken
in the analysis is the most appropriate because it distinguishes daily bi-modal
(e.g. crepuscular) patterns that might be important but would not be detected by
insolation alone (i.e. insolation at dawn or dusk would be similar values but
occur at different times of day that might be ecologically important).
Additionally, as the reviewer points out, results from the analysis and haul-out
predictions are more intuitive and applicable to aerial survey methods when 
using hour-of-day.
:::

> ***Line 278:** Please help your reader understand what is meant by your term
> "horizontal shift".  It is clear that you have a graphical representation in
> your mind that has not been shared with your reader. Please rewrite to avoid
> "jargon".*

> ***Line 285:** Please help your reader understand what could be "important"
> about these terms. Please write for clarity, avoiding jargon.*

> ***Figure 4:** Please use standard conventions for plot titles.  Your title
> deserves a statement in your Results or Discussion section.  It does not
> follow the expectations for a publication figure title.*

::: author-response
**\[author response\]** We have removed plot titles and edited the figure
captions to include the title information to be consistent with the journal
styling and requirements. We've also revised many of the
figure legends to better communicate the information to the reader and ensure
if any figures are presented out of context (e.g. without the figure caption) in
other venues the information is clear.
:::

> ***Figure 4:** Please scale your color ramp to enable better visualization of
> the subject: limit the color ramp to the modeled values (which appear to be
> from zero to \~0.5).*

::: author-response
**\[author response\]** We have adjusted the color ramp such that the pattern
for bearded seals is more visible. We've also noted this change/difference in
the figure caption so the reader is aware.
:::

> ***Figure 4:** Please switch to a color ramp that is suitable for readers with
> color perception disabilities such as the viridis color ramps.*

@@

::: author-response
**\[author response\]** The color ramp used has been updated to use palettes
from the R package 'scico' have been tested and confirmed to be colorblind
friendly and appropriately balanced for scientific interpretation. We initially
used the viridis color palette for these plots but the 'scico' palette colors
provided additional visual detail we feel is important.
:::

> ***Figure 4:** This is unclear. Your plot indicates a y-axis of local solar
> hour which varies from 0 to 24. This is clearly not fixed at local noon.*

@@

::: author-response
**\[author response\]** This was an error and has been fixed. We thank the
reviewer for recognizing this and highlighting the mistake.
:::

> ***Line 323:** Please use standard English. "Ontongeny" is generally
> understood to describe the physical development of an individual organism.
> Here you are referencing the onset of of behaviors within the life of an
> organism. It may benefit your readers to use a term that is more direct, such
> as "development".*

@@

::: author-response
**\[author response\]** We have replaced 'ontongeny' with 'development' as
suggested
:::

> ***Line 326-327:** Given that Ribbon seals occupy a high latitude region
> during this time and given that there is scarcely any twilight during this
> season, please consider using a term other than "crepuscular".*

@@

::: author-response
**\[author response\]** Subsequent model runs (with improved QA/QC on telemetry
data) have reduced the indication of any crepuscular or bi-modal haul-out
behavior of ribbon seals and we no longer feel it is worth mentioning in the
paper so it has been removed.
:::

> ***Line 331-332:** This is a fair summary of your modeling efforts. However,
> it raises the questions as to what these seals are responding to. Do they
> sense air pressure and make a decision on their haulout behavior or are low
> pressure conditions associated with low insolation and they are responding to
> that?  Given that you can readily extract insolation from the NARR, please
> consider use of insolation as a covariate that may more directly relate to
> seal behavior.*

@@

::: author-response
**\[author response\]** See previous responses regarding the potential use of
'insolation' from NARR or other reanalysis products.
:::

> ***Figure 5:** Please use standard figure titling and not use a statement that
> should below in a PowerPoint slide or in your Results and Discussion section.*
>
> ***Figure 6:** Please use standard conventions for plot titles. Your title
> deserves a statement in your Results or Discussion section. It does not
> follow the expectations for a publication figure title.*

@@

::: author-response
**\[author response\]** See previous response regarding our choice in plot
titles and subsequent edits
:::

> ***Figure 8:** Please label your plots using standard capitalization and
> words, instead of all capitalized abbreviations.*

@@

::: author-response
**\[author response\]** We have made edits based on this suggestion
:::

> ***Line 360-362:** This belongs in the discussion, not results.*

::: author-response
**\[author response\]** The suggestion to move the mention that there "were
still some years where the pattern shown was informed by a small number of
individuals that may not represent population-level patterns" to the discussion
is noted and changes have been made.
:::

> ***Line 371-372:** Please indicate the season (or months of the year) in this
> statement.*

@@

::: author-response
**\[author response\]** Noted and edited to include mention of the spring season
:::

> ***Line 375-376:** Do you mean to state that "haulout behavior modeled from
> telemetered haulout behavior records may represent population-level behavior
> patterns."*

::: author-response
**\[author response\]** Noted and edited accordingly
:::

> ***Line 377-378:** Feltz and Fay (1966) offer (an amazingly cool) functional
> explanation of the thermal constraints on hair follicle growth.  Whether seals
> during specific study times are optimizing hair follicle growth is unknown.*

@@

::: author-response
**\[author response\]** We have edited this statement for additional clarity
:::

> ***Line 418-419:** Yes, and you have had access to direct estimates of
> insolation from the NARR.  Can you incorporate that and find a tighter model
> of haulout behavior.  This would be an improvement for design and
> interpretation of aerial survey efforts.*

@@

::: author-response
**\[author response\]** See previous comments regarding the potential use of
insolation values from reanalysis products
:::

> ***Line 420-421:** You may state this if you first offered an estimate of
> haulout behavior and confidence intervals on that behavior state, then
> compared that with your estimated haulout behavior modeled with the use of
> co-variates. Once you perform this, you may state that inclusion of weather
> co-variates improved haulout state estimates. However, you have not done
> this. If you wish to state this, please do this.*

@@

::: author-response
**\[author response\]** Here we simply state that 'our results highlight the
importance of incorporating weather covariates' and that importance is indicated
with the reported 'type III' F statistics for the weather covariate terms in the
results section. We are clear in our methods that the use of pseudo-likelihoods
in the analysis prevent us from conducting typical model selection with AIC or
similar as suggested. We have moved a separate discussion of comparison with
previous analysis results here to provide additional comparative context.
:::

> ***Line 425:** This term suggests some evaluation of a threshold above which
> haulout behavior shifts. Please write for clarity and for what you have
> presented in your results.*

@@

::: author-response
**\[author response\]** We have edited this sentence for improved clarity
:::

> ***Line 455:** Please be clear as to why you anticipate greater unmodeled
> heterogeneity of availability for Arctic seals? Here you have considered
> Arctic seals, have you not? Or are you using a different definition of the
> Arctic other than the ice bearing waters occupied by ice-associated seals?*

::: author-response
**\[author response\]** We have edited this statement to now read:

The most challenging element in developing
availability correction factors is with annual variability. It can be difficult
to get a sufficient sample size to estimate year-specific correction factors,
particularly because research teams would likely need to deploy bio-loggers on
seals and conduct aerial surveys concurrently, requiring considerably more
personnel and money. One possible suggestion is to include year as a random
effect within models for aerial survey counts such that, without specific
knowledge of any particular year, the among-year variance is included in the
modeled standard errors. Regardless of the specific approach, future estimates
of Arctic seal abundance will require specific consideration of annual
variability and changes in the timing of peak haul-out behavior when estimating
trends, as one will not know if moderate differences in abundance estimates are
attributable to changes in abundance or changes in haul-out behavior.

The interpreted emphasis on 'Arctic seals' was not intended. Instead, the key
point is that there is likely unmodeled heterogeneity related to the seasonal
distribution and annual timing of peak haul-out behavior. This is an
acknowledgement that our model is not a complete picture or perfect
representation of seal haul-out behavior. In an environment like the Arctic that
is undergoing dramatic change and inter-annual variation in climate and sea-ice
distribution, this is important to keep in mind.
:::

> ***Line 459-464:** Here you provide the discussion required to make the
> statement that inclusion of weather and other covariates results in a
> different estimate of haulout proportion. Can you extend this to indicate an
> improvement in the confidence interval on this estimated proportion?*

@@

::: author-response
**\[author response\]** Noted and this section has been moved to that discussion
:::

> ***Line 465-466:** Please be specific. Here you linked the telemetered haulout
> behavior to available covariate data then modeled haulout behavior.*

@@

::: author-response
**\[author response\]** Noted and additional specifics have been added to this
statement
:::

> ***Line 467-473:** This can be stated in the introduction so that you can
> clearly indicate the framing of your study. "We modeled availability from
> telemetered haulout behavior records of seals that haulout on ice where they
> may be viewed by aerial surveys (list the species considered).*

@@

::: author-response
**\[author response\]** This section has been moved to the introduction.
:::

## Reviewer 2

> *One suggestion I would share with the authors is the addition of a
> limitations section in the discussion. The limitations were outlined
> throughout the methods, but a succinct accounting of these limitations in the
> discussion would allow readers who do not want to carefully dissect all of the
> modeling in the methods to be able to better interpret the robustness of the
> results and conclusions.*

::: author-response
**\[author response\]** We thank the reviewer for this question and, in response
to this and other reviewer feedback, have moved much of the limitations outlined
in the methods to the Discussion section and hope this provides a clearer
picture to the reader regarding limits of our analysis.
:::

> *I still had one remaining question on why the authors couldn't run the models
> with and without the sea ice concentration as a predictor. The comparative
> results might be interesting (I understand why the correction factor might be
> biased...but given that this is a baseline study, it would have been an
> interesting comparison to see whether it was or wasn't a decent predictor for
> the different seal species).*

::: author-response
**\[author response\]** We think we've addressed this in edits and comments to
reviewer #1. We think the inclusion of sea ice as a predictor is better suited
for a habitat use/habitat preference study and we hope to pursue that in future
analyses. We have also emphasized that our reliance on flipper-mounted
bio-loggers for a portion of our data limited our understanding of movement
and use away from sea-ice.
:::

### 2024-02-18 analysis/paper/Letter_Response_to_Reviewers.qmd

> *The authors need to explain more how they are attributing statistical
> significance. For example, in lines 344-345, they ascribe strong influence of
> temperature with a p-value of 0.064.\
> \
> Not to be one who is fixated on a p-value of 0.05 denoting significance, I was
> a bit surprised to see the same authors, in lines 364-366, to state that there
> is 'no indication that the observed trend is meaningful'for spotted seal adult
> females having an R2 of 0.767, p=0.022. If the extent of sea ice explains
> nearly 80% of the variance in peak haul-out probability, how is it that the
> author's call this a 'trend'? This requires more explanation. As does the
> statement in line 427-428.*

::: author-response
**\[author response\]** We appreciate the reviewer highlighting this confusion
and apologize for it. These were typos or mis-coded values that snuck into the
section. We have reviewed and revised all of them to ensure they are correct and
that we are consistent in our attribution of significance. We have tried to be
judicious in our use of the work 'significant'.
:::

> *I found 463-464 confusing. This is an important part of the discussion (using
> your framework to assess previous results from another study), but it could be
> more clearly written. The sentence, 'Applying models that ignore age, sex, and
> year effects...under the current analysis framework' is very confusing and
> doesn't seem to follow the argument of the paper easily. Could this be
> reworded so that the full intent of the paragraph is communicated?*

::: author-response
**\[author response\]** Reviewer #1 also commented on this section and offered
suggestions for editing. With both comments in mind, we have made edits to the
text and hope this, now, more clearly communicates how our results compare with
previous approaches.
:::

> *I appreciate the acknowledgement that it is difficult to know whether
> differences in abundance estimates are attributable to changes in abundance or
> changes in haul-out behavior and the potential proposed solution (lines
> 450-458) Has anyone done a smaller-scale study that would address this
> inherent problem?*

::: author-response
**\[author response\]** To our knowledge a smaller-scale study addressing this
issue has not been done
:::

@@

## Reviewer 3 (M. Carter)

> *I have only a few minor comments, mostly where the text would benefit from
> extra clarity:\
> *
>
> ***L43:** Word missing? ...Bearing \[and\] Chukchi Seas...*\

@@

::: author-response
**\[author response\]** noted and fixed
:::

> **L63:** It may be useful to highlight somewhere that matching seal haulout
> behaviour to environmental conditions with any degree of certainty requires
> location data as well as wet-dry data, and that modern bio-logging devices can
> collect both data streams simultaneously.\

@@

::: author-response
**\[author response\]** thank you for this suggestion and we have edited this
paragraph to include mention regarding the ability of modern bio-logging devices
to collect both behavioral and location data streams which allows inclusion of
environmental covariates.
:::

> **L82-83:** All citations given here relate to ice-associated seals, but these
> techniques are also relevant to seals that haul-out on land. Lonergan et al.
> (2013) Aquatic Conservation: Marine and Freshwater Ecosystems 23: 135-144 is a
> key example from non-polar regions.\

::: author-response
**\[author response\]** we have expanded the list of references to include
Lonergan et al
:::

> **L110:** Clarify why meteorological factors affecting walruses may be
> relevant to the study species.\

@@

::: author-response
**\[author response\]** thank you for the suggestion. We've edited the sentence
to, now, read:

"Previously estimated availability correction factors accounted for variables
such as time-of-day and day-of-year but did not investigate the impact of
meteorological variables. Such variables have been shown to influence walrus
haul-out behavior and we expect weather conditions to also influence seal
haul-out behavior and including them within the model framework will benefit our
estimates of seal availability during aerial surveys."
:::

> **L120-124:** A map of the study area with deployment areas indicated would be
> useful for those readers not familiar with the local geography.\

@@

> **L126-128:** More information is required on the location data type (e.g. GPS
> or ARGOS?) and data retrieval (e.g. archival loggers or ARGOS-relay?).\

::: author-response
**\[author response\]** we have added the following sentence:

"All data were transmitted by the deployed instruments via the Argos satellite
network and location data were either derived from Argos transmissions or
transmitted FastLoc GPS data."
:::

> **L133-135:** Please clarify, were there any cases where only
> head/back-mounted tags were deployed? If so, how did you deal with the
> potential for false haul-out records due to prolonged surface activity?\

@@

::: author-response
**\[author response\]** we've clarified the language in this paragraph so it's
clear that deployments consisted of head/back tag only, flipper tag only, and
combined. Examination of the haul-out records hasn't indicated there's a
propensity for these tags to record extended dry periods when surface resting at
sea for the head/back tags.
:::

> **L135:** At this stage you have not yet explained what type of data each
> device collects, so we do not know what the hourly percent-dry observations
> from the flipper tag were preferred to.\

@@

::: author-response
**\[author response\]** thank you for noting this. we have moved the later
paragraph (starting line 165) up so it is before this section.
:::

> **L139:** Which species?\

::: author-response
**\[author response\]** we have removed the 'for some species' from this
sentence
:::

> **L136-144:** This could be clearer. I assume the "some species" part refers
> only to pelage, but it appears at first read to refer to the whole list of
> metrics. I would suggest re-wording this section to improve clarity regarding
> which factors were used for which species and with what degree of certainty
> age class could be determined.\

@@

::: author-response
**\[author response\]** as noted in the previous response, we've removed the
'for some species' and think this improves the clarity of this section.
Unfortunately, we don't have any way to quantify a degree of certainty for age
class determination across the various field teams responsible for field
identification.
:::

> **L146:** How many YOY were advanced to subadult?\

::: author-response
**\[author response\]** there were 11 YOY advanced to subadult. We considered
adding this into the results but struggled to find a place where the information
fit and was relevant.
:::

@@

> **L188:** On land as well as ice?\

::: author-response
**\[author response\]** we have edited this sentence to state 'also haul out on
land'
:::

> **L194:** It is not clear to me if all tags for a given species were deployed
> in the same region, or across multiple regions (a map of the study area with
> deployment regions per species would help here). Given regional differences in
> phenology mentioned in the Introduction (L52-53), did you consider whether
> region should be included as a covariate?\

@@

::: author-response
**\[author response\]** we have added a figure/map as suggested
:::

> **Figs 4, 6, 8:** How is that hour of the day is fixed at local solar noon if
> predictions are shown for all hours? Also, is it possible to show uncertainty
> (lower and upper CIs) for these figures, perhaps in Supplementary Materials?\

@@

::: author-response
**\[author response\]** noted and removed 'also'
:::

> **L341:** Word missing? ...and \[spent\] more time out of the water...\

::: author-response
**\[author response\]** noted and we've made the suggested correction
:::

> **L387-390:** It is not clear what species was studied by Thometz et al, and
> thus maybe a citation is needed for the information relating to ribbon and
> spotted seals.\

@@

::: author-response
**\[author response\]** we've added a clarification that the Thometz et al study
and data presented was for bearded seals.
:::

> **L393:** Use "both species" instead of "all species" to avoid confusion.\

::: author-response
**\[author response\]** we've edited to specifically state 'ribbon and spotted
seals' to avoid any confusion
:::

> \
> The paper could benefit from a brief concluding paragraph summarising the
> importance of the work in the context of climate change and population trends,
> as outlined in the Introduction.
